---
title: "Introduction to Binomial Package"
author: "shefali sastry"
date: "`4/30/2019`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(binomial)
```

# Binomial Package

## An Introduction to the Binomial Random Variable
 
This Binomial package calculates probabilities of a Binomial random variable because it allows us to calculate the probabilities to find the number of successes in a fixed number of random trials performed under identical conditions (assuming a constant probability of success on each trial).


## What this package contains

This specific package contains 15 functions that the user can use, 10 of which are main functions, and 5 of which are method functions that support these main functions due to their plotting capabilities. Here is a list of what is provided:


* Main Functions
    * `bin_choose()`
    * `bin_probability()`
    * `bin_distribution()`, Method: `plot.bindis()`
    * `bin_cumulative()`, Method: `plot.bincum()`
    * `bin_var()`, Method: `print.binvar()`
    * Summary/Auxiliary functions: `bin_mean()`, `bin_variance()`, `bin_mode()`, `bin_skewness()`, `bin_kurtosis()`
    
* Method Functions
    * `plot.bindis()`, 
    * `plot.bincum()`
    * `print.binvar()`
    * `summary.binvar()`
    * `print.summary.binvar()`
```{r echo=FALSE}
## Check Functions
# These functions will not be called by the user as they are primarily used to check for valid values.
# Function check_prob()
#This function tests whether an input prob is a valid probability value
check_prob <- function(prob) {
  if (prob > 1 | prob < 0) {
    stop('p has to be a number betwen 0 and 1')
  }
  else {
    return(TRUE)
  }
}

# Function check_trials()
# This function tests whether an input trials is a valid value for a number of trials (i.e. n is a non-negative integer)
check_trials <- function(trials){
  if (trials <0) {
    stop('invalid trials value')
  } else {
    return(TRUE)
  }
}

# Function check_success()
# This function tests the success input is a valid value for number of successes
check_success <- function(success, trials){
  if(is.vector(success) == TRUE & sum(success <= trials) == length(success))
    return(TRUE)
  else
    stop('invalid success value')
}
```

## How to Apply These Functions

### bin_choose()
###### This function calculates the number of combinations in which k successes can occur in n trials.
```{r}
bin_choose(n = 5, k = 2)
bin_choose(5, 0)
```


### bin_probability()
###### This function calculates the probability of getting k successes in n trials assuming a certain probability of success.
```{r}
# probability of getting 2 successes in 5 trials
# (assuming prob of success = 0.5)
bin_probability(success = 2, trials = 5, prob = 0.5)

# 55 heads in 100 tosses of a loaded coin with 45% chance of heads
bin_probability(success = 55, trials = 100, prob = 0.45)
```

### bin_distribution()
###### This function calculates the distribution of the probabilities for each number of successes.
```{r}
# binomial probability distribution
bin_distribution(trials = 5, prob = 0.5)
```

### bin_cumulative()
###### This function calculates the distribution of the probability of getting at most k number of successes in n trials.

```{r}
# binomial cumulative distribution
bin_cumulative(trials = 5, prob = 0.5)
```


### bin_var()
###### This function calculates a list of the  binomial variables by inputting n trials and at a probability level.
```{r}
bin1 <- bin_variable(trials = 10, p = 0.3)
bin1
```

```{r echo=FALSE}
aux_mean <- function(trials,prob) {
  multiply_trials_prob <- trials*prob
  return(multiply_trials_prob)
}
  
bin_mean <- function(trials, prob){
  check_trials(trials)
  check_prob(prob)
  return(aux_mean(trials, prob))
}

aux_variance <- function(trials,prob) {
  variance_trials_prob <- trials*prob*(1-prob)
  return(variance_trials_prob)
}

bin_variance <- function(trials, prob){
  check_trials(trials)
  check_prob(prob)
  return(aux_variance(trials, prob))
}

aux_mode <- function(trials,prob) {
  mode_trials_prob <- floor(trials*prob + prob)
  return(mode_trials_prob)
}

bin_mode <- function(trials, prob){
  check_trials(trials)
  check_prob(prob)
  return(aux_mode(trials, prob))
}

aux_skewness <- function(trials,prob) {
  skewness_trials_prob <- (1-(2*prob))/(sqrt(trials*prob*(1-prob)))
  return(skewness_trials_prob)
}

bin_skewness <- function(trials, prob){
  check_trials(trials)
  check_prob(prob)
  return(aux_skewness(trials, prob))
}

aux_kurtosis <- function(trials,prob) {
  kurtosis_trials_prob <- ((1-6*prob)*(1-prob))/((trials*prob))*(1-prob)
  return(kurtosis_trials_prob)
}

bin_kurtosis <- function(trials, prob){
  check_trials(trials)
  check_prob(prob)
  return(aux_kurtosis(trials, prob))
}

```
## Summary Functions
### bin_mean()
###### This calculates the mean of a binomial distribution.
```{r}
bin_mean(10, 0.3)
```

### bin_variance()
###### This calculates the variance of a binomial distribution.
```{r}
bin_variance(10, 0.3)
```

### bin_mode()
###### This calculates the variance of a binomial distribution.
```{r}
bin_mode(10, 0.3)
```

### bin_skewness()
###### This calculates the skewness, or a measure of the asymmetry of the probability distribution of a random variable of a binomial distribution.
```{r}
bin_skewness(10, 0.3)
```

### bin_kurtosis()
###### This calculates the skewness, or a measure of the “tailedness” of the probability distribution of a random variable of a binomial distribution.
```{r}
bin_kurtosis(10, 0.3)
```





## Method Functions
### plot.bindis()
###### This function creates a barplot the distribution of the probabilities for each number of successes.
```{r}
plot.bindis <- function(yes){
  barplot(yes$probability, names.arg = yes$success, xlab = "successes", ylab = "probability")
}

# plotting binomial probability distribution
dis1 <- bin_distribution(trials = 5, prob = 0.5)
plot(dis1)

```


### plot.bincum()
###### This function creates a line plot of the cumulative distribution of the probabilities for each number of successes.
```{r}
plot.bincum <- function(cum1){
  plot(cum1$success, cum1$cumulative, type = "o")
  lines(cum1$success, cum1$cumulative)
}

dis2 <- bin_cumulative(trials = 5, prob = 0.5)
plot(dis2)
```


### print.binvar()
###### This function prints the content of a list of the  binomial variables by inputting n trials and at a probability level.
```{r}
print.binvar <- function(x){
  cat('"Binomial Variable"\n\n')
  cat("Parameters\n", append = TRUE)
  cat("-number of trials:", x$trials, append = TRUE)
  cat("\n-probability:", x$prob, append = TRUE)
}
bin1 <- bin_variable(trials = 10, p = 0.3)
bin1
```


### summary.binvar()
###### This function contains a full summary description of an object "binvar". It includes the elements: trials, mean, variance, mode, skewness, kurtosis.
```{r}
summary.binvar <- function(x,...){
  object1 <- list(trials = x$trials, prob = x$prob, mean = aux_mean(x$trials, x$prob), variance = aux_variance(x$trials, x$prob), mode = aux_mode(x$trials, x$prob), skewness = aux_skewness(x$trials, x$prob), kurtosis = aux_kurtosis(x$trials, x$prob) )
  class(object1) = "summary.binvar"

  return(object1)
}
bin_summary <- bin_variable(trials = 10, p = 0.3)
bin_summary1 <- summary(bin_summary)
bin_summary1
```

### print.summary.binvar()
###### This function prints the content and includes a full summary description of "summary.binvar".
```{r}
print.summary.binvar <- function(x){
  cat('"Summary Binomial"\n\n')
  cat("Parameters\n", append = TRUE)
  cat("-number of trials:", x$trials, append = TRUE)
  cat("\n-prob of success:", x$prob, append = TRUE)
  cat('\n\n"Measures"')
  cat("\n-mean:", aux_mean(x$trials, x$prob), append = TRUE)
  cat("\n-variance:", aux_variance(x$trials, x$prob), append = TRUE)
  cat("\n-mode:", aux_mode(x$trials, x$prob), append = TRUE)
  cat("\n-skewness:", aux_skewness(x$trials, x$prob), append = TRUE)
  cat("\n-kurtosis:", aux_kurtosis(x$trials, x$prob), append = TRUE)
}

bin1 <- bin_variable(trials = 10, p = 0.3)
binsum1 <- summary(bin1)
binsum1

```
